---
title: "Comparing Coffee Characteristics by Country of Origin"
subtitle: "*Using tmap and ShinyApp*"
author: "Prasiddhi Thapaliya"
title-slide-attributes:
  data-slide-number: none
format: revealjs
editor: visual
execute:
  echo: false
---

## Research Question

#### Do coffees grown in certain countries/regions produce different flavor profiles?

I made an interactive Shinyapp map that allows users to choose from these five flavor traits:

-   Flavor

-   Acidity

-   Sweetness

-   Aroma

-   Total Cup Points (total quality score from 1-100)

## The Data

##### scrapped and cleaned by Github user @jldbc

-   Three .csv files with consumer reviews from the Coffee Quality Institute

    -   Arabica, Robusta, both (merged)

-   Information on

    -   quality measures

    -   bean data (such as processing method),

    -   farm/origin data (such as altitude)

-   Reviews were mix of qualitative and quantitative, converted to numeric scale

## The Data

```{r importing data, echo = FALSE, include = FALSE}
library(tidyverse)
library(rmarkdown)
arabica <- read_csv("data/arabica_data_cleaned.csv")
robusta <- read_csv("data/robusta_data_cleaned.csv")
merged <- read_csv("data/merged_data_cleaned.csv")
```

```{r glimpse average, echo = T}
arabica %>%
  select(Country.of.Origin, Acidity, Aroma, Flavor, Sweetness, Total.Cup.Points)
```

## Top Ten (Arabica)

::: panel-tabset
### Acidity

```{r}
arabica %>%
  select(Acidity, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Acidity)
```

### Aroma

```{r}
arabica %>%
  select(Aroma, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Aroma)
```

### Flavor

```{r}
arabica %>%
  select(Flavor, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Flavor)
```

### Sweetness

```{r}
arabica %>%
  select(Sweetness, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Sweetness)
```

### Total

```{r}
arabica %>%
  select(Total.Cup.Points, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Total.Cup.Points)
```
:::

## Top Ten (Robusta)

::: panel-tabset
### Acidity

```{r}
robusta %>%
  select(Salt...Acid, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Salt...Acid)
```

### Aroma

```{r}
robusta %>%
  select(Fragrance...Aroma, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Fragrance...Aroma)
```

### Flavor

```{r}
robusta %>%
  select(Flavor, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Flavor)
```

### Sweetness

```{r}
robusta %>%
  select(Bitter...Sweet, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Bitter...Sweet)
```

### Total

```{r}
robusta %>%
  select(Total.Cup.Points, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Total.Cup.Points)
```
:::

## Top Ten (Both)

::: panel-tabset
### Acidity

```{r}
merged %>%
  select(Acidity, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Acidity)
```

### Aroma

```{r}
merged %>%
  select(Aroma, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Aroma)
```

### Flavor

```{r}
merged %>%
  select(Flavor, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Flavor)
```

### Sweetness

```{r}
merged %>%
  select(Sweetness, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Sweetness)
```

### Total

```{r}
merged %>%
  select(Total.Cup.Points, Country.of.Origin) %>%
  slice_max(n = 10, order_by = Total.Cup.Points)
```
:::

## Limitations

-   Much more data on arabica than robusta species

-   Some countries have many more observations than others

    -   ex. Ethiopia vs. Panama
    -   Using mean -\> sensitive to outliers

-   Correlation vs Causation

    -   Two coffees grown at different altitudes, different countries

    -   Is it *really* country of origin that creates the flavor profile?

# ShinyApp!

## Predictive Modelling

#### Can we predict region of origin based on flavor profile?

-   Supervised learning algorithm

-   Grouping each country by region (using "World" dataset)

-   Addition to Shiny dashboard

    -   Insert values for flavor profile, guess region of origin?

# To be Continued...
